### 2020-02-20 bioclim_data.Rmd
Bioclim

---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("raster")
install.packages("sp")
```

```{r}
library("tidyverse")
library("raster")
library("sp")
```

```{r}
#Download worldclim data 
clim = getData('worldclim', var='bio', res=2.5) 
gain(clim) = 0.1
```

```{r}
#look up SpatialPoints 
# https://gis.stackexchange.com/questions/227585/using-r-to-extract-data-from-worldclim
```