### 2020-02-20 bioclim_data.Rmd
Bioclim

---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("raster")
install.packages("sp")
```

```{r}
library("tidyverse")
library("raster")
library("sp")
```

```{r}
#Download worldclim data 
clim = getData('worldclim', var='bio', res=2.5) 
gain(clim) = 0.1 #values in bioclim dataset are scaled by 10
```

```{r}
# https://gis.stackexchange.com/questions/227585/using-r-to-extract-data-from-worldclim

coords <- data.frame(x=c(-80.2, -80.4, -80.9, -81.8, -81.5, -80, -78.1, -78.4, -77.1, -77.4, -74.9, -74.1, 127.8, 130.6, 130.6, 130.3, 131.5, 135.8, 136, 136.7, 139, 140.9),
                     y=c(26.67, 27.58, 28.85, 30.33, 31.13, 33.75, 35.03, 36.35, 37.05, 38.63, 39.77, 40.73, 26.2, 31.6, 32.65, 33.25, 34.18, 35, 35.35, 36.56, 37.91, 38.29))
points <- SpatialPoints(coords)
values <- extract(clim,points)

```
The bioclimate variables used were: minimum temperature of coldest month