### 2020-04-06 _albopictus_cpp_MS.Rmd
analysis and files for Apr1 draft

---
title: "Ae. albopictus CPP manuscript"
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
output:
  html_document: default
  word_document: default
  pdf_document: default
bibliography: MyLibrary.bib
---

```{r eval=FALSE, message=FALSE, include=FALSE}
install.packages("tidyverse") #1.2.1
install.packages("drc") #3.0.1
install.packages("stringr")
install.packages("lubridate")
install.packages(c("maps", "mapdata"))
devtools::install_github("dkahle/ggmap")
install.packages("scatterpie")
install.packages("ggedit")
install.packages("ggrepel")
install.packages("papeR")
install.packages("cowplot")
install.packages("flextable")
install.packages("stats")
install.packages("ggbiplot")
```

```{r message=FALSE, echo=FALSE, include=FALSE}
library("papeR")
library("flextable")
library("drc") #3.0.1
library("knitr") #1.25
library("tidyverse") #1.2.1
library("stringr") #1.4.0
library("geosphere") #1.5.10
library("lubridate") #1.7.4
library("maps")
library("mapdata")
library("ggedit")
library("ggrepel")
library("cowplot")
library("ggrepel")
library("stats")
library("ggbiplot")

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300)
```

```{r datasheet, echo=FALSE, include=FALSE}
all_data<-data.frame(read.csv("CPP_experiment_data_collated.csv"))
#Filter out empty/discarded oviposition papers 
blockAB<-dplyr::filter(all_data, TotalLarvaeEmbryos>0)
```

```{r yields, echo=FALSE, include=FALSE}
#Calculate yield/fraction of initial count
blockAB$yield <- blockAB$TotalLarvaeEmbryos/blockAB$InitialEggs 
hist(blockAB$yield, main="Yield (final/initial count) per oviposition paper")
plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos)
```

```{r filter papers, echo=FALSE, include=FALSE}
#removing papers with <60% or >100% initial count
blockAB_unfiltered<-blockAB
blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
```

```{r summarize, echo=FALSE, include=FALSE}
DI_summarize <- function(blockAB) {
DI_summary <- blockAB %>% 
  group_by(CageID) %>% 
  dplyr::summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>%   filter(h1+h2+emb>100)
DI_summary$total<-DI_summary$h1+DI_summary$h2+DI_summary$emb
DI_summary$DI<-DI_summary$emb/DI_summary$total
return(DI_summary)
}
```

```{r, echo=FALSE, include=FALSE}
DI_summary<-DI_summarize(blockAB_filtered)
```

```{r totals_plot, echo=FALSE, include=FALSE}
hist(DI_summary$total, main="Total number of viable eggs per cage")
```

```{r DI_plot, echo=FALSE, include=FALSE}
#FIG S2
#plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK")
DI_summary$pop<-factor(DI_summary$pop, levels=c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK"))
DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop, fill=country)) + geom_point(shape=21) + facet_wrap(~fct_reorder(pop, lat)) + theme_bw() + theme(legend.position="NULL", text=element_text(size=18)) + 
xlab("Photoperiod (hours)") + ylab("Dipause incidence") + scale_fill_manual(values=c("black","gray")) 
```

```{r CPP_calc_2018, echo=FALSE, include=FALSE}
#From Kevin Emerson
# the following fits the models of the given  using data from the
# data.frame d separately for the various treatments using a 5 parameter
# logistic model
d.model.fit <- drm(DI ~ pp, data = DI_summary, curveid = pop, fct = LL.5())
#plot(d.model.fit, log = "")
# output the parameters of the model fits:
summary(d.model.fit)
# estimate the 50% intercept
ED(d.model.fit, 50)
cpp_values<-data.frame(ED(d.model.fit, 50))

pm<-drc:::predict.drc(d.model.fit, type="response")
```

```{r CPP_calc_2008, echo=FALSE, include=FALSE}
DI_summary_2008<-data.frame(read.csv("CPP_2008.csv"))
d.model.fit.2008 <- drm(DI ~ photoperiod, data = DI_summary_2008, curveid = pop, fct = LL.5())

#plot(d.model.fit.2008, log = "")
# output the parameters of the model fits:
summary(d.model.fit)
# estimate the 50% intercept
ED(d.model.fit.2008, 50)
cpp_values_2008<-data.frame(ED(d.model.fit.2008, 50))
```

```{r collate_cpp_values, echo=FALSE, include=FALSE}
cpp_2018<-data.frame(ED(d.model.fit, 50))
cpp_2018<-rownames_to_column(cpp_2018, var="pop") 
cpp_2018$pop<-str_sub(cpp_2018$pop,3,5)
popdata_2018<-DI_summary %>% group_by(pop) %>% 
  dplyr::summarize(lat=first(lat), country=first(country))
cpp_2018<-right_join(cpp_2018, popdata_2018, by="pop")
cpp_2018$year<-rep("2018", nrow(cpp_2018))
colnames(cpp_2018)[colnames(cpp_2018)=="Std..Error"] <- "StdError"
cpp_2008<-data.frame(ED(d.model.fit.2008, 50)) %>% rownames_to_column(var="pop")
cpp_2008$pop<-str_sub(cpp_2008$pop,3,5)
colnames(cpp_2008)[colnames(cpp_2008)=="Std..Error"] <- "StdError"
popdata_2008<-DI_summary_2008 %>% group_by(pop) %>% 
  dplyr::summarize(country=first(country), lat=first(lat), year=first(year)) 
cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")
#1988 data not used in present study but can load it anyway
cpp_1988<-data.frame(read.csv("CPP_1988.csv")) 
cpp_1988$pop<-na_if(cpp_1988$pop, "NA")
cpp_1988$StdError<-na_if(cpp_1988$StdError, "NA")

#2008 and 2018 data
cpp_all<-rbind(cpp_2008, cpp_2018) 
write.csv(cpp_all, "cpp_calcs_filtered.csv")
```

```{r}
#FIG S1
yield_plot<-ggplot(blockAB, aes(x=TotalLarvaeEmbryos, y=yield)) + geom_point(alpha=0.5) + xlab("Number of viable eggs\n on oviposition paper") + ylab("Viable eggs/initial egg count on paper") + theme_bw() + theme(text=element_text(size=18))

cpp_filtered<-read.csv("cpp_calcs_filtered.csv") %>% dplyr::select(pop, Estimate, country, lat, year) %>% filter(year==2018)
cpp_unfiltered<-read.csv("cpp_calcs_unfiltered.csv") %>% dplyr::select(pop, Estimate, country, lat, year) %>% filter(year==2018)
names(cpp_filtered)[2]<-"CPP_est_filt"
names(cpp_unfiltered)[2]<-"CPP_est_unfilt"
cpp_compare<-left_join(cpp_filtered,cpp_unfiltered,by="pop")

filter_compare_plot<-ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0) + xlab("CPP without filtering criteria") + ylab("CPP calculated with filtering criteria") + theme_bw() + theme(text=element_text(size=18))

quality_control_fig<-plot_grid(yield_plot, filter_compare_plot, labels="AUTO")
```

```{r cppmodel, echo=FALSE, include=FALSE}
cppmodel<-lm(Estimate~lat*country*year, data=cpp_all)
cppmodel_result<-data.frame(anova(cppmodel))
```

```{r, echo=FALSE, include=FALSE}
#basic breakdown
ggplot(cpp_all, aes(x=lat, y=Estimate, color=country)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  facet_grid(~year) + stat_smooth(method=lm, se=FALSE)
```

```{r, echo=FALSE, include=FALSE}
#FIG 2: CPP by latitude, year and country
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008
cpp_08_18_shifts_bw <- ggplot(cpp_all, aes(x=lat, y=Estimate, fill=country, shape=year)) +
  geom_point(size=4, stroke=1) + scale_fill_manual(values=c("black", "gray"), name="Country") +
  scale_shape_manual(values=c(22,21), name="Year") +
  stat_smooth(method=lm, se=FALSE, aes(linetype=year, color=country), size=1.5) + scale_color_manual(values=c("black", "gray"), name="Country") +  scale_linetype_manual(values=c("longdash","solid"), name="Year") + 
  theme(legend.position=c(.8,.25), text=element_text(size=24),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlab("Latitude") + ylab("Critical photoperiod (h)") + 
  guides(linetype=guide_legend(override.aes=list(color="black"), legend.key.width=unit(5,"line")))

```

```{r incidence_lat_year, echo=FALSE, include=FALSE}
DI_short_2008<-filter(DI_summary_2008, photoperiod==8)
DI_short_2018<-filter(DI_summary, pp==8) %>% group_by(pop) %>% 
  dplyr::summarize(country=first(country), lat=first(lat), photoperiod=first(pp), emb=sum(emb), total=sum(total))
DI_short_2018$DI<-DI_short_2018$emb/DI_short_2018$total
DI_short_2018<-DI_short_2018[,c(1:4,7)] 
DI_short_2018$year<-rep("2018",nrow(DI_short_2018))
DI_short<-rbind(DI_short_2008, DI_short_2018)
write_csv(DI_short, "diapause_incidence_8h_2008_2018.csv")
```

```{r dimodel, echo=FALSE, include=FALSE}
#DI by lat/country/year comparison
dimodel_poly_us<-lm(DI~lat^2 + lat, data=filter(DI_short, country=="US"))
dimodel_lat_us<-lm(DI~lat, data=filter(DI_short, country=="US"))
dimodel<-lm(DI~lat*country*year, data=DI_short)
dimodel_result<-data.frame(anova(dimodel))
#checking for normality of residuals
dimodel.stdres<-rstandard(dimodel)
qqnorm(dimodel.stdres, ylab="Standardized Residuals", xlab="Normal Scores", main="Normal Probability Plot")
qqline(dimodel.stdres)

#Spearman rank correlation coefficients
cor.test(filter(DI_short_2, country=="US", year==2018)$lat, filter(DI_short, country=="US", year==2018)$DI, alternative="two.sided", method="spearman")
cor.test(filter(DI_short_2, country=="JP", year==2018)$lat, filter(DI_short, country=="JP", year==2018)$DI, alternative="two.sided", method="spearman")

#comparison of latitude-matched DIs in each country between 2008 and 2018
#exact latitude match in US; matched by rank in JP
wilcox.test((filter(DI_short, country=="US", year==2018, pop!="PBS") %>% arrange(lat))$DI, (filter(DI_short, country=="US", year==2008) %>% arrange(lat))$DI, paired=TRUE)
wilcox.test((filter(DI_short, country=="JP", year==2018) %>% arrange(lat))$DI, (filter(DI_short, country=="JP", year==2008) %>% arrange(lat))$DI, paired=TRUE)

#FIG S3
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008
country_DI_bw<-ggplot(DI_short, aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("gray","black"), name="Country") + scale_shape_manual(values=c(22,21), name="Year") + geom_point(size=3) +   theme(legend.position=c(.8,.25), text=element_text(size=18),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
xlab("Latitude") + ylab("Short-day diapause incidence") + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("longdash","solid"), name="Year") + scale_color_manual(values=c("gray","black"), name="Country")

```

```{r maps, echo=FALSE, include=FALSE}
usa<-map_data("usa")
jp<-map_data("japan")
pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))
us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="blue") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-85, -65)) +
  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw() 
japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="red") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  + xlim(c(125, 147)) +
  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())

#bw locations map
us_map_bw<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 2) + xlim(c(-87, -65)) + ylim(c(24,48)) +
  geom_label_repel(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge_x=1.5, size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm")) + xlab("") + ylab("")

japan_map_bw<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 2)  + xlim(c(125, 147)) + ylim(c(24,48)) +
  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm")) + xlab("") + ylab("Latitude")

map_panels_bw<-plot_grid(japan_map_bw, us_map_bw)
```

```{r}
#date of first first
#first frost data
#importing and reshaping table
ff_us <- read.csv("FF_US.csv", check.names=FALSE) %>% gather(year, DFF, "1998":"2018", factor_key=FALSE)
ff_us$year<-as.numeric(ff_us$year)
ff_jp <- read.csv("FF_JP.csv", check.names=FALSE) %>% gather(year, DFF, "1998":"2018", factor_key=FALSE)
ff_jp$year<-as.numeric(ff_jp$year)
#adding latitudes
locations<-read.csv("mapping_pops_cpp_expt.csv")
ff_us<-left_join(ff_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
ff_jp<-left_join(ff_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
ff_all<-rbind(ff_us, ff_jp)
ff_all$pop<-as.factor(ff_all$pop)
ff_all<-mutate(filter(ff_all, year!=2018), interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
                                                  year<2008 ~ 'y1998-2007'))

ff_interval_summaries<-ff_all %>% group_by(pop, interval) %>% dplyr::summarise(dff_mean=mean(DFF), dff_sd=sd(DFF), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))

ff_all_summary<-ff_all %>% group_by(pop) %>% dplyr::summarise(dff_mean=mean(DFF), dff_sd=sd(DFF), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))

#FIG 1B
#black=Japan, gray=US 
DFF_decade<-ggplot(filter(ff_interval_summaries, only_2008=="n", interval=="y2008-2017"), aes(x=latitude, y=dff_mean, fill=country)) + geom_point(size=3, shape=21, stroke=1) + scale_fill_manual(values=c("black", "gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") +
  theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +  scale_linetype_manual(values=c("dotted","solid"), name="Country") + geom_errorbar(aes(ymin=dff_mean-dff_sd, ymax=dff_mean+dff_sd)) + xlab("Latitude") + ylab("Date of first frost (ordinal day)")

```

```{r}
locations_dff<-plot_grid(map_panels_bw, DFF_decade, nrow=2, labels="AUTO")
locations_dff
```

```{r, echo=FALSE, include=FALSE}
#GDD 400 deadline data
#importing and reshaping table
gdd400_us <- read.csv("GDD400_US.csv", check.names=FALSE) %>% gather(year, gdd400, "1998":"2018", factor_key=FALSE)
gdd400_us$year<-as.numeric(gdd400_us$year)
gdd400_jp <- read.csv("GDD400_JP.csv", check.names=FALSE) %>% gather(year, gdd400, "1998":"2018", factor_key=FALSE)
gdd400_jp$year<-as.numeric(gdd400_jp$year)
#adding latitudes
locations<-read.csv("mapping_pops_cpp_expt.csv")
gdd400_us<-left_join(gdd400_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
gdd400_jp<-left_join(gdd400_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
gdd400_all<-rbind(gdd400_us, gdd400_jp)
gdd400_all$pop<-as.factor(gdd400_all$pop)
gdd400_all<-mutate(gdd400_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
                                                  year<2008 ~ 'y1998-2007'))

gdd465_us <- read.csv("GDD465_US.csv", check.names=FALSE) %>% gather(year, gdd465, "1998":"2018", factor_key=FALSE)
gdd465_us$year<-as.numeric(gdd465_us$year)
gdd465_jp <- read.csv("GDD465_JP.csv", check.names=FALSE) %>% gather(year, gdd465, "1998":"2018", factor_key=FALSE)
gdd465_jp$year<-as.numeric(gdd465_jp$year)
locations<-read.csv("mapping_pops_cpp_expt.csv")
gdd465_us<-left_join(gdd465_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
gdd465_jp<-left_join(gdd465_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
gdd465_all<-rbind(gdd465_us, gdd465_jp)
gdd465_all$pop<-as.factor(gdd465_all$pop)
gdd465_all<-mutate(gdd465_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
                                                  year<2008 ~ 'y1998-2007'))
gdd_all<-left_join(gdd400_all, gdd465_all)
gdd_all$gdd<-(gdd_all$gdd400 + gdd_all$gdd465)/2

```

```{r, echo=FALSE, include=FALSE}
#FIG S4
gdd_deadline_plots<-ggplot(filter(gdd465_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd465, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21, size=3) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + 
theme(legend.position="bottom", text=element_text(size=18)) 
```

```{r echo=FALSE, include=FALSE}
#adding 2008 and 2018 CPP values 
cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% dplyr::rename("CPP_YEAR"="year")
gdd465_all_new<-left_join(gdd465_all, cpp_year, by=c("pop"))
gdd465_all_new$CPP_YEAR<-as.factor(gdd465_all_new$CPP_YEAR)

gdd_all_new<-left_join(gdd_all, cpp_year, by="pop")
gdd_all_new$CPP_YEAR<-as.factor(gdd_all_new$CPP_YEAR)

```

```{r echo=FALSE, include=FALSE}
#GDD deadlines averaged across all years
gdd_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop) %>% dplyr::summarise(gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))

lat_gdd_average<-ggplot(filter(gdd_summaries, only_2008=="n"), aes(x=latitude, y=gdd465_mean, fill=country)) + geom_point(shape=21) + scale_fill_manual(values=c("black", "gray")) + geom_errorbar(aes(ymin=gdd465_mean-gdd465_sd, ymax=gdd465_mean+gdd465_sd, color=country)) + scale_color_manual(values=c("black","gray")) + theme_bw() + theme(legend.position="bottom") + xlab("Latitude (N)") + ylab("GDD Deadline (ordinal day)")

```

```{r echo=FALSE, include=FALSE}
#plotting GDD deadlines and CPP
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008

gdd_08 <- ggplot(filter(gdd465_all_new, year==2007, CPP_YEAR==2008), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=22) + stat_smooth(method=lm, se=FALSE, aes(color=country), linetype="longdash") + scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
    xlab("GDD Deadline 2007 (ordinal day)") + ylab("CPP (h), 2008") + ylim(12.3,14.2)

gdd_18 <- ggplot(filter(gdd465_all_new, year==2017, CPP_YEAR==2018), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
xlab("GDD Deadline 2017 (ordinal day)") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)

```

```{r echo=FALSE, include=FALSE}
#plotting CPP by GDD deadlines averaged over intervals
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008

interval_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop, interval, CPP_YEAR) %>% dplyr::summarise(gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), cpp=mean(Estimate), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))

gdd_cpp_interval_18<-ggplot(filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018), aes(x=gdd465_mean, y=cpp, fill=country)) + geom_point(size=4, stroke=1, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") +   theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlab("Mean GDD deadline 2008-2017 \n(ordinal day)") + ylab("CPP, 2018 (h)") + ylim(12.3, 14.2)

gdd_cpp_interval_08<-ggplot(filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008), aes(x=gdd465_mean, y=cpp, fill=country)) + geom_point(size=4, stroke=1, shape=22) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country), linetype="longdash") + scale_color_manual(values=c("black", "gray"), name="Country") +   theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlab("Mean GDD deadline 1998-2007 \n(ordinal day)") + ylab("CPP, 2008 (h)") + ylim(12.3, 14.2)

```

```{r echo=FALSE, include=FALSE}
#FIG 4
gdd_cpp_plot<-plot_grid(gdd_08, gdd_18, gdd_cpp_interval_08, gdd_cpp_interval_18, nrow=2, labels="AUTO")

```

```{r echo=FALSE, warning=FALSE}
#FIG S5A
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008

gdd_by_interval<-ggplot(filter(interval_summaries, CPP_YEAR==2018), aes(x=latitude, y=gdd465_mean, fill=country, shape=interval)) + 
  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(22,21), name="Period") + geom_errorbar(aes(ymin=gdd465_mean-gdd465_sd, ymax=gdd465_mean+gdd465_sd, color=country)) +
  stat_smooth(method=lm, se=FALSE, aes(linetype=interval, color=country)) + scale_linetype_manual(values=c("longdash","solid"), name="Period") + 
  scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position=c(.8,.8), text=element_text(size=16),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlab("Latitude") + ylab("Mean GDD Deadline")

```

```{r echo=FALSE, include=FALSE}
data_18<-gdd465_all_new %>% filter(CPP_YEAR==2018) %>% filter(year==2017)
model1<-lm(Estimate~gdd465*country, data=data_18)

data_08<-gdd465_all_new %>% filter(CPP_YEAR==2008) %>% filter(year==2007)
model2<-lm(Estimate~gdd465*country, data=data_08)

model3<-lm(cpp~gdd465_mean*country, data=filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018))
model4<-lm(cpp~gdd465_mean*country, data=filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008))

anova(model1)
anova(model2)
anova(model3)
anova(model4)

```

```{r echo=FALSE, include=FALSE}
#2015 Diapause Incidence Field Trial
field_data<-read.csv("DI_2015_FIELD.csv")
field_data$WeekDate <- reorder(field_data$WeekDate, field_data$Week)

#field eggs only
field.model.fit <- drm(DI ~ DOY, data = field_data, fct = LL.5())
summary(field.model.fit)
# estimate the 50% intercept
cpp_field<-data.frame(ED(field.model.fit, 50))

field_data_plot<-ggplot(field_data, aes(x=DOY, y=DI)) + geom_point(size=3, shape=21, fill="gray") + 
  theme(legend.position=c(.8,.25), text=element_text(size=24),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
xlab("Ordinal day (2015)") + ylab("Diapause incidence") + geom_vline(aes(xintercept=cpp_field$Estimate), linetype="dotted") + coord_cartesian(xlim=c(193,365))

#2014 GDD diagram
man2014<-read.csv("2014_Mean_Temp_MAN.csv") %>% gather(degreetype, degrees, "CountDegrees":"UncountDegrees")

man2014_gdd_plot<-ggplot(filter(man2014, DOY_2014>193), aes(x=DOY_2014, y=degrees, fill=degreetype)) +
    theme(text=element_text(size=24), legend.position="none",
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  geom_bar(stat="identity", width=1) + 
  scale_fill_manual(values=c("dark gray","light gray")) + 
  geom_hline(yintercept=10, linetype="dashed") + geom_vline(xintercept=251) +
  xlab("Ordinal Day (2014)") + ylab("Mean Daily Temp (C)")

#Fig3
field_and_gdd2014<-plot_grid(field_data_plot, man2014_gdd_plot, nrow=2, align="hv", labels="AUTO")

```

```{r}
#climate variation data
tmean_us<-read.csv("Monthly_Tmean_US.csv", check.names=FALSE) %>% gather(year, meantemp, "1998":"2018", factor_key=FALSE)
#In order to look at winter temperatures, need to combine Nov/Dec with Jan/Feb/March of the following year. January and February are reassigned in this table to the previous year to create seasonal years
tmean_us$year<-as.numeric(tmean_us$year) 
tmean_us$seasonalyear<-if_else(tmean_us$month %in% c("Jan","Feb","Mar"), tmean_us$year-1, tmean_us$year)

#make a list of the coldest months by seasonal year and location
coldest_months_US <- tmean_us %>% group_by(pop, seasonalyear) %>% slice(which.min(meantemp)) %>% filter(!(seasonalyear %in% c(1997,2018))) %>% select(pop, seasonalyear, month)

tmin_us<-read.csv("AvgTmin_US.csv", check.names=FALSE) %>% gather(year, mintemp, "1998":"2018", factor_key=FALSE)
tmin_us$year<-as.numeric(tmin_us$year) 
tmin_us$seasonalyear<-if_else(tmin_us$month %in% c("Jan","Feb","Mar"), tmin_us$year-1, tmin_us$year) 
tmin_us<-filter(tmin_us, !(seasonalyear %in% c(1997,2018)))

tmax_us<-read.csv("AvgTmax_US.csv", check.names=FALSE) %>% gather(year, maxtemp, "1998":"2018", factor_key=FALSE)
tmax_us$year<-as.numeric(tmax_us$year) 
tmax_us$seasonalyear<-if_else(tmax_us$month %in% c("Jan","Feb","Mar"), tmax_us$year-1, tmax_us$year) 
tmax_us<-filter(tmax_us, !(seasonalyear %in% c(1997,2018)))

#make a table of the min, mean, and max temperature of the coldest month of each seasonal year in each location
all_temps_us<-left_join(tmin_us, tmean_us)
all_temps_us<-left_join(all_temps_us, tmax_us)
all_temps_us<-select(all_temps_us, pop, month, seasonalyear, mintemp, meantemp, maxtemp)
cold_month_temps_us<-inner_join(all_temps_us, coldest_months_US)
cold_month_temps_us$country=rep("US",nrow(cold_month_temps_us))

#same process for Japan
tmean_jp<-read.csv("Monthly_Tmean_JP.csv", check.names=FALSE) %>% gather(year, meantemp, "1998":"2018", factor_key=FALSE)
#In order to look at winter temperatures, need to combine Nov/Dec with Jan/Feb/March of the following year. January and February are reassigned in this table to the previous year to create seasonal years
tmean_jp$year<-as.numeric(tmean_jp$year) 
tmean_jp$seasonalyear<-if_else(tmean_jp$month %in% c("Jan","Feb","Mar"), tmean_jp$year-1, tmean_jp$year)

#make a list of the coldest months by seasonal year and location
coldest_months_jp <- tmean_jp %>% group_by(pop, seasonalyear) %>% slice(which.min(meantemp)) %>% filter(!(seasonalyear %in% c(1997,2018))) %>% select(pop, seasonalyear, month)

tmin_jp<-read.csv("AvgTmin_JP.csv", check.names=FALSE) %>% gather(year, mintemp, "1998":"2018", factor_key=FALSE)
tmin_jp$year<-as.numeric(tmin_jp$year) 
tmin_jp$seasonalyear<-if_else(tmin_jp$month %in% c("Jan","Feb","Mar"), tmin_jp$year-1, tmin_jp$year) 
tmin_jp<-filter(tmin_jp, !(seasonalyear %in% c(1997,2018)))

tmax_jp<-read.csv("AvgTmax_JP.csv", check.names=FALSE) %>% gather(year, maxtemp, "1998":"2018", factor_key=FALSE)
tmax_jp$year<-as.numeric(tmax_jp$year) 
tmax_jp$seasonalyear<-if_else(tmax_jp$month %in% c("Jan","Feb","Mar"), tmax_jp$year-1, tmax_jp$year) 
tmax_jp<-filter(tmax_jp, !(seasonalyear %in% c(1997,2018)))

#make a table of the min, mean, and max temperature of the coldest month of each seasonal year in each location
all_temps_jp<-left_join(tmin_jp, tmean_jp)
all_temps_jp<-left_join(all_temps_jp, tmax_jp)
all_temps_jp<-select(all_temps_jp, pop, month, seasonalyear, mintemp, meantemp, maxtemp)
cold_month_temps_jp<-inner_join(all_temps_jp, coldest_months_jp)
cold_month_temps_jp$country=rep("JP",nrow(cold_month_temps_jp))

#add latitude information and combine US/Japan
locations<-read.csv("mapping_pops_cpp_expt.csv")
cold_month_temps_us<-left_join(cold_month_temps_us, dplyr::select(locations, pop, latitude, only_2008), by=c("pop"))
cold_month_temps_jp<-left_join(cold_month_temps_jp, dplyr::select(locations, pop, latitude, only_2008), by=c("pop"))
cold_months_all<-rbind(cold_month_temps_us, cold_month_temps_jp)

cold_months_all<-mutate(cold_months_all, interval=case_when(seasonalyear<=2017 & seasonalyear>=2008 ~ 'y2008-2017', seasonalyear<2008 ~ 'y1998-2007'))

#calculate annual range and annual standard deviation
#start with monthly means
us_var <- tmean_us %>% filter(!(seasonalyear %in% c(1997,2018))) %>% group_by(pop, seasonalyear) %>%
  dplyr::summarise(tempannualrange=max(meantemp)-min(meantemp), tempseasonality=sd(meantemp))
jp_var <- tmean_jp %>% filter(!(seasonalyear %in% c(1997,2018))) %>% group_by(pop, seasonalyear) %>%
  dplyr::summarise(tempannualrange=max(meantemp)-min(meantemp), tempseasonality=sd(meantemp))
all_var<-rbind(us_var, jp_var)
all_var<-left_join(all_var, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))

all_var<-mutate(all_var, interval=case_when(seasonalyear<=2017 & seasonalyear>=2008 ~ 'y2008-2017', seasonalyear<2008 ~ 'y1998-2007'))

tf_us <- read.csv("TotalFrostDays_US.csv", check.names=FALSE) %>% gather(year, TotalFrostDays, "1998":"2018", factor_key=FALSE)
tf_us$year<-as.numeric(tf_us$year)

tf_jp <- read.csv("TotalFrostDays_JP.csv", check.names=FALSE) %>% gather(year, TotalFrostDays, "1998":"2018", factor_key=FALSE)
tf_jp$year<-as.numeric(tf_jp$year)
#adding latitudes
locations<-read.csv("mapping_pops_cpp_expt.csv")
tf_us<-left_join(tf_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
tf_jp<-left_join(tf_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
tf_all<-rbind(tf_us, tf_jp)
tf_all$pop<-as.factor(tf_all$pop)
tf_all<-mutate(filter(tf_all, year!=2018), interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
                                                  year<2008 ~ 'y1998-2007'))

frost_all<-select(left_join(ff_all, tf_all), pop, latitude, country, year, DFF, TotalFrostDays, only_2008, interval)
names(frost_all)[names(frost_all) == 'year']<-'seasonalyear'

clim_vars_all<-left_join(cold_months_all, all_var)
names(clim_vars_all)[names(clim_vars_all) == 'month'] <- 'coldest_month'
names(clim_vars_all)[names(clim_vars_all) == 'mintemp'] <- 'mintemp_coldest_month'
names(clim_vars_all)[names(clim_vars_all) == 'meantemp'] <- 'meantemp_coldest_month'
names(clim_vars_all)[names(clim_vars_all) == 'maxtemp'] <- 'maxtemp_coldest_month'
clim_vars_all<-select(clim_vars_all, pop, latitude, country, seasonalyear, coldest_month, mintemp_coldest_month, meantemp_coldest_month, maxtemp_coldest_month, tempannualrange, tempseasonality, interval)

clim_vars_all<-left_join(clim_vars_all, frost_all)

cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% dplyr::rename("CPP_YEAR"="year")

clim_vars_all<-left_join(clim_vars_all, cpp_year, by=c("pop"))
clim_vars_all$CPP_YEAR<-as.factor(clim_vars_all$CPP_YEAR)

```

```{r}
cp<-dplyr::select(filter(clim_vars_all, CPP_YEAR==2018), mintemp_coldest_month, meantemp_coldest_month, maxtemp_coldest_month, tempannualrange, tempseasonality, DFF, TotalFrostDays)
names(cp)<-c("ColdestMonthMin","ColdestMonthMean","ColdestMonthMax","TempAnnualRange","TempSeasonality","DateFirstFrost","TotalFrostDays")
cv<-dplyr::select(filter(clim_vars_all, CPP_YEAR==2018), pop, latitude, country, seasonalyear, interval)

```

```{r}
clim.pca <- prcomp(cp,
                 center = TRUE,
                 scale. = TRUE)
pca_by_country<-ggbiplot(clim.pca, groups=cv$country) + 
        theme(text=element_text(size=16), legend.position=c(.8,.8),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlim(values=c(-3,3)) + ylim(values=c(-3,3)) +
  scale_color_manual(name="Country", values=c("black","gray"))

pca_by_interval<-ggbiplot(clim.pca, groups=cv$interval) + 
        theme(text=element_text(size=16), legend.position=c(.8,.8),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlim(values=c(-3,3)) + ylim(values=c(-3,3)) +
  scale_color_manual(name="Interval", values=c("black","gray"))

pca_by_year<-ggbiplot(clim.pca, groups=cv$seasonalyear) + 
        theme(text=element_text(size=18), legend.position="bottom",
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlim(values=c(-3,3)) + ylim(values=c(-3,3)) +
  scale_color_gradient(name="Year")

pca_by_latitude<-ggbiplot(clim.pca, groups=cv$latitude) +         
  theme(text=element_text(size=18), legend.position="bottom",
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlim(values=c(-3,3)) + ylim(values=c(-3,3)) +
  scale_color_gradient(name="Latitude")

pcas_for_fig<-plot_grid(pca_by_country, pca_by_interval, nrow=1)
```

```{r}
#FIG S5
plot_grid(gdd_by_interval, pcas_for_fig, nrow=2, labels="AUTO")
```